---
title: "Homework 9"
author: "Carolyn Hanrahan"
date: "2023-03-29"
output: html_document
---
# Question 1: 
Use the code that you worked on in Homework #8 (creating fake data sets), and re-organize it following the principles of structured programming. Do all the work in a single chunk in your R markdown file, just as if you were writing a single R script. Start with all of your annotated functions, preliminary calls, and global variables. The program body should be only a few lines of code that call the appropriate functions and run them in the correct order. Make sure that the output from one function serves as the input to the next. You can either daisy-chain the functions or write separate lines of code to hold elements in temporary variables and pass them along.

## Loading Libraries: 
```{r}
library(sqldf)
library(dplyr)
library(tidyverse)
```


# Creation of global variables
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data(ToothGrowth)
```

# Source Files and Run Workflow 

```{r}
setwd("C:/Users/Carolyn/Desktop/Computational Biology/HanrahanBio381")

source("Functions_Homework9.R")

get_data(ToothGrowth) 

x <- calculate_mean(ToothGrowth)

y <- calculate_sd(ToothGrowth)

FakeDataFrame <- create_fake_dataset(ToothGrowth)

pivot_dataset_longer(ToothGrowth)

create_boxplot(ToothGrowth)

calculate_ANOVA(ToothGrowth)

```
- For question 1, I created multiple functions in my Functions_Homework9.R script. These functions brought in the data, calculated mean and standard deviation of the data, created a boxplot and finally calculated the ANOVA. I used similar code to what I wrote for homework 8, but had to alter it slightly to make it more generic such that it fit into a function. It was important that I assigned variables x and y in my Rmarkdown file in order for the preceding functions to run properly. This lab dealt a lot with local vs. global variables, and figuring out when and where to specify variables was challenging. Hopefully this document has some semblance of the right approach.  

# Question 2
Once your code is up and working, modify your program to do something else: record a new summary variable, code a new statistical analysis, or create a different set of random variables or output graph. Do not rewrite any of your existing functions. Instead, copy them, rename them, and then modify them to do new things. Once your new functions are written, add some more lines of program code, calling a mixture of your previous functions and your new functions to get the job done.


```{r}
get_data(rnorm)

PracticeData <- rnorm(50, mean=20, sd=2)
calculate_ANOVA(PracticeData)
create_boxplot2(PracticeData)

```

- For question 2, I altered the create boxplot function to create boxplots that are green. I then created a new random dataset using the rnorm function, and passed this new data through my functions. Lastly, I calculated the ANOVA for my new random dataset. Thank you and sorry this is a boring answer to question 2!