---
title: "Homework 5"
author: "Carolyn Hanrahan"
date: "2023-02-15"
output: html_document
---

## Question 1

Assign to the variable n_dims a single random integer between 3 and 10.

```{r}
n_dims <- runif(n=1, min=3, max=10)
print(n_dims)
```
  - Here I used the runif (random uniform distribution) function to pick a random number between 3 and 10 and assign it to the variable.

Create a vector of consecutive integers from 1 to n_dims2
```{r}
vector1 <- c(1:n_dims^2)
print(vector1)
```
  - Here I used the concatenate function to create a vector with values from 1 to n_dims squared. 
  
Use the sample function to randomly reshuffle these values.
```{r}
shuffle_vector <- sample(x=vector1)

```
  - I decided to call the shuffled version of this vector "shuffle_vector"

create a square matrix with these elements.
```{r}
matrix1 <- matrix(data=shuffle_vector, nrow=n_dims, ncol=n_dims, byrow=FALSE)

```
   - Because n_dims will change when it is regenerated each time, the matrix row and column were assigned as the variable n_dims so the matrix will change, as well. 


print out the matrix.

```{r}
print(matrix1)

```

find a function in r to transpose the matrix.

```{r}
transpose_matrix <- t(matrix1)

```
  - Searched for "transpose" function on Google!

print it out again and note how it has changed.

```{r}
print(transpose_matrix)

```

calculate the sum and the mean of the elements in the first row and then the last row.

```{r}
print(transpose_matrix)
```
sum: 
```{r}

sum <- sum(transpose_matrix[1, ] + transpose_matrix[n_dims, ])

print(sum)

```
mean: 
```{r}
mean <- mean(transpose_matrix[1, ] + transpose_matrix[n_dims, ]) 
print(mean)
```
 - Note that I used the variable n_dims as the last row of the vector in the above computations because the number of rows and columns will shift when new n_dims values are computed. 
 
 

read about the eigen() function and use it on your matrix


```{r}
?eigen()

eigen <- eigen(transpose_matrix)
print(eigen)

```

look carefully at the elements of $values and $vectors in the output. What kind of numbers are these?

   - these are complex numbers! They have imaginary numbers as indicated by the i.

dig in with the typeof() function to figure out their type.

```{r}
eigen$values
typeof(eigen$values)
```

if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.


## Question 2 
Create a list with the following named elements:

my_matrix, which is a 4 x 4 matrix filled with random uniform values
my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_letters, which is a 26-element vector of all the lower-case letters in random order.
```{r}
runif <- runif(n=16, min=0, max= 100)
my_matrix <- matrix(data=runif, nrow=4, ncol=4)

print(my_matrix)

```
```{r}
my_other_vector <- runif(n=100, min=0, max=100)

my_logical <- my_other_vector >45 

print(my_logical)
```
```{r}
my_letters <- letters
print(my_letters)

```


Then, complete the following steps:

create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.

```{r}
my_list <- list(my_matrix[2,2], my_logical[2], my_letters[2])

print(my_list)
```

use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(my_list[[1]])
typeof(my_list[[2]])
typeof(my_list[[3]])
```

combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}

atomic_vector <-c(my_list[[1]],my_list[[2]],my_list[[3]])
print(atomic_vector)

```
  - In the above code, I called upon the values in position 1, 2, and 3 in the list, and concatenated them into a vector.
  
what is the data type of this vector?

```{r}
typeof(atomic_vector)
```
## Question 3 
Create a data frame with the two variables (= columns) and 26 cases (= rows) below:


call the first variable my_unis and fill it with 26 random uniform values from 0 to 10

```{r}
my_unis <- runif(26, min=0, max=10)
```
call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
my_letters2 <- sample(LETTERS)
print(my_letters2)
```

  - I used "LETTERS" which denotes all 26 letters in their capitalized form. 

```{r}
data_frame <- data.frame(my_unis, my_letters2)
print(data_frame)
```

for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.

```{r}
print(data_frame)

data_frame[sample(1:26, 4), 1] <- NA

data_frame

```
  - Because there are 26 rows, you can randomly sample 1:26 and it will select rows for you. I specified 4 in my code to indicate that I am looking to sample 4 rows, and used the <-NA to assign those four rows the values NA. Note that ( ,1) means every row, column number 1).


for the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
is.na(data_frame)
```

re-order the entire data frame to arrange the second variable in alphabetical order

```{r}

data_frame[order(my_letters2), ]


reorder_data_frame <-data_frame[order(my_letters2), ]

```

```{r}
print(reorder_data_frame)

```
calculate the column mean for the first variable.

```{r}
mean(data_frame[ ,1], na.rm= TRUE)
data_frame[ ,1]

```
  - to calculate the column mean for the first variable, I used the mean function, referencing our data frame, and all of the rows in it. I specifically had to say na.rm = TRUE in order to remove the Null values. Having null values in our calculation would cause issues. Thank you :)


