---
title: "Homework 12"
author: "Carolyn Hanrahan"
date: "April 19, 2023"
output: html_document
---

# Homework 11
  - For homework 11 I plan on using the ToothGrowth built-in dataset in R to practice making some fun new ggplot graphics and base R plots. I begin by creating some basic plots and work on increasingly improving my data visualiazation through multiple iterations. 
  
  
## Importing libraries and data: 


```{r}
library(ggplot2)
library(ggthemes)
library(patchwork)
data("ToothGrowth")
head(ToothGrowth)

d <- ToothGrowth
```
## Some Practice Graphs 

```{r}

hist(d$len) 
table(d$len)
plot(table(d$len))

```

## Basic Boxplot

```{r}
e <- boxplot(d$len~d$dose)

```



## Utilizing ggplot :)


```{r}

p1 <- ggplot(data=d) +
               aes(x=dose,y=len)+
               geom_point()

print(p1)

```
  - Adding more complexity...

```{r}
p2 <- p1 + theme_solarized()
print(p2)
```


  - and more...
```{r}
p3 <- ggplot(data=d) +
  aes(x=dose,y=len) + 
  geom_point(size=7,
             shape=24,
             color="black",
             fill="steelblue") +
  labs(title="Vitamin C Treatment",
       subtitle="vitamin c given in units of 0.5, 1, and 2 mg",
       x="Vitamin C treatment (mg)",
       y="Tooth Length (mm)") +
  xlim(0,2) + ylim(0,40)

print(p3)
```

  - Testing out some totally new stuff from the R graph gallery: Ridge plot
  
```{r}

# library
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)


x <- ggplot(data=d, aes(x = len, y = dose, fill = dose, group = dose)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "left")
  labs(title='Tooth Length by Vitamin C Treatment Group')


print(x)



```

  - Progressively adding more to the ridge plot...
  
  
```{r}

y <- ggplot(d, aes(x = len, y = dose, fill = dose, group = dose)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Dose (mg)", option = "F") +
  labs(title = 'Tooth Length by Vitamin C Dosage')+
  labs(caption="A graph created by Carolyn Hanrahan on 4/19/23")

print(y)

ggsave("ToothLengthRidgeGraph.pdf")  #saving my graph to a pdf.



```

```{r}
z<- ggplot(d, aes(x = len, y = dose)) + 
  geom_density_ridges(scale = 1) + facet_wrap(~dose)


print(z)

```


  - this graph puts each plot into its own little section. 


## A New type of plot: 

  - 3D plot!This plot is not very well suited to this data because we only have 2 numeric categories. I am going to keep the code here, though:
  
```{r}
library(rgl)



plot3d <- plot3d( 
  x=d$dose, y=d$len, z=d$supp, 
  col = "red","blue","yellow", 
  type = 's', 
  radius = .1,
  xlab="Dose", ylab="Tooth Length", zlab="Sup")

print(plot3d)




```

## Examining Supplement Type and Tooth Length (Orange juice vs.Ascorbic Acid)

```{r}

boxplot(d$len~d$supp)


s <- ggplot(data=d) +
  aes(x=supp,y=len,fill=supp) + 
  geom_boxplot() + 
  theme(legend.position="none")+
  plot_annotation(title="Tooth Length vs.Supplement Type",caption="Graph made by Carolyn Hanrahan on 4/19/23")

print(s)


```

  - I feel as though this is a very nice and informative graph. You can clearly see that the supplement type of OJ result in greater tooth length. 

## Saving my Graph 

  - I saved my last graph and named it "s"

```{r}

ggsave("SupplementType.pdf")

```
## Conclusion 
  - Overall I really enjoyed exploring the different graphing options in this lab!I feel like the last boxplot and some of the ridgeline graphs are effective examples of decent data visualizations. 

