---
title: "Homework10"
author: "Carolyn Hanrahan"
date: "2023-04-05"
output: html_document
---
# loading libraries 
```{r setup, include=FALSE}
library(ggplot2)
library(tidyr)
library(tidyverse)
library(dplyr)
```

# Question 1
  Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.
  


# if statement within a for loop
```{r}
z <-c(-2,0,0,1,2,3,5)

counter <- 0

for(i in 1:length(z)) {        
  if(z[i] == 0) {counter<-counter + 1}
   }


```

  
# Question 2
  Use subsetting instead of a loop to rewrite the function as a single line of code.
  
  Below is a failed example of me attempting this problem. This answer technically works but is not general enough to run new vectors through. 
```{r}
x <- c(0,0,2,3,1,2)
counter2 <- 0

#counter2 <- if(x[1]|x[2]|x[3]|x[4]|x[5]|x[6]==0){counter2<-counter2+1}


```
  Below is a more generalized way of subsetting: 
  
```{r}
y <- c(0,0,2,1,2,1,2,1) #random vector 

counter2 <-sum(y==0)

```



# Question 3
  Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
  
```{r}
row <- 2
col <- 2

#######################################
# FUNCTION: make matrix
# packages: none
# purpose: create a new matrix (see output)
# input: number of rows and columns 
# output: matrix of these dimensions in which each element is the product of the row number times the column number 
#-------------------------------------- 
makematrix <- function(row, col) {
 
  m <- matrix(nrow=row,ncol=col)
  
  for(i in 1:row){
  
    for(j in 1:col){
   
       m[i,j] <- i*j
    
  }  
    
  }
  
 return(m)
print(m)      

}

makematrix(8,2)

```


# Question 4: 
  In the next few lectures, you will learn how to do a randomization test on your data. We will complete some of the steps today to practice calling custom functions within a for loop. Use the code from the March 31st lecture (Randomization Tests) to complete the following steps:
  
  a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable. 
 
 My first attempt(not correct..): 
```{r}
# control <- rnorm(20,mean=5,sd=1)
# treatment1<-rnorm(20,mean=15,sd=2)
# treatment2<-rnorm(20,mean=25,sd=2)
#
# dataframe <- data.frame(control,treatment1,treatment2)
#
# print(dataframe)
#
# pivotdf <- pivot_longer(dataframe,cols=c('treatment1'))
#
# print(pivotdf)
```
  Using Nick's code: 
```{r}
trt_group <- c(rep("Control",4),rep("Treatment1",4),rep("Treatment2",4))
print(trt_group)

z <- c(runif(4) + 1, runif(4) + 10, runif(4)+2)
print(z)

df <- data.frame(trt=trt_group,res=z)
print(df)


```

  b. Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.
  
```{r}
#######################################
# FUNCTION: re-shuffle
# packages: none
# purpose: 
# input:
# output: 
#-------------------------------------- 
shuffle_data <- function(df) {
                
  df[,2] <- sample(df[,2])
    
          
            mean <- df%>%          
            group_by(trt) %>% 
            summarize(Mean=mean(res)) 
                
            return(mean)

            print(mean)
 
}
              
shuffle_data(df)           

```


c. 
  Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.

```{r}
#dataframe <- data.frame(repNumber=numeric(0),ControlMean=numeric(0),Treatment1Mean=numeric(0),Treatment2Mean=numeric(0))

dataframe<-data.frame(matrix(NA, nrow=100, ncol=4))

colnames(dataframe) <- c('replicateNumber','ControlMean','Treatment1Mean','Treatment2Mean')

print(dataframe)
print(df)
 
for(i in 1:100) {
 x<-shuffle_data(df)
  
  
   dataframe[i,1] <-i
   dataframe[i,2] <- x[1,2] 
   dataframe[i,3] <- x[2,2] 
   dataframe[i,4] <- x[3,2]
   
   }
 
print(dataframe)
dataframe$replicateNumber
```

d. 
  Use qplot() to create a histogram of the means for each reshuffled group. Or,if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?

```{r}

hist(dataframe$ControlMean)

hist(dataframe$Treatment1Mean)

hist(dataframe$Treatment2Mean)

```


  - Note that the histograms appear different, which is what we are looking for between the different treatments. 




